Actual Causes as Changes of State in Continuous Time

### 2021-01-21 docs/resources/CogSci_paper/paper.tex

Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}. Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{}, force dynamics \cite{wolff} and physical simulation~\cite{gerstenberg}.  Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.  On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.

Most accounts of actual causation are consistent with one of two broad views of causation: the counterfactual approach and the physical process approach. The counterfactual approach suggests that the tripping of relay Q29BD is a cause of the elevator stopping because if the relay had not tripped then the elevator would not have stopped. The physical process view suggests that the relay tripping is a cause of the elevator stopping because the two are linked by a physical process involving the transmission of force or energy.  Both accounts need additional machinery in order to specify which among the many causes of an event is singled out as the main cause. Our continuity account fits most naturally with the physical process approach, and can be viewed as an attempt to bring out some implications of this approach for actual causation.

@@

\nocite{halpernp01}

>>>>

Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}. Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.  Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.  On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.

Most accounts of actual causation are consistent with one of two broad views of causation: the counterfactual approach and the physical process approach~\cite{hall04}. The counterfactual approach suggests that the tripping of relay Q29BD is a cause of the elevator stopping because if the relay had not tripped then the elevator would not have stopped. The physical process view suggests that the relay tripping is a cause of the elevator stopping because the two are linked by a physical process involving the transmission of force or energy.  Both accounts need additional machinery in order to specify which among the many causes of an event is singled out as the main cause. Our continuity account fits most naturally with the physical process approach, and can be viewed as an attempt to bring out some implications of this approach for actual causation.

@@
